Was Devon Allen Unjustly Disqualified at the 2022 World Track and Field Championships?

### 2023-04-04 Manuscript/manuscript.tex
amsfonts needed for mathbb

\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{cleveref}
\hypersetup{colorlinks = true, linkcolor = blue, citecolor=blue, urlcolor = blue}
\usepackage{natbib}
\usepackage{float}
\usepackage{setspace}
\usepackage{pdfpages}
\usepackage[pagewise]{lineno}
\usepackage{mwe}
\usepackage{comment}
%\linenumbers*[1]
% %% patches to make lineno work better with amsmath
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
 \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
 \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
 \renewenvironment{#1}%
 {\linenomath\csname old#1\endcsname}%
 {\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
 \patchAmsMathEnvironmentForLineno{#1}%
 \patchAmsMathEnvironmentForLineno{#1*}}%

@@

2) and 3)
Outside the blog posts from LetsRun.com, there have not been any published
papers regarding Devon Allen's disqualification in particular.  This is not
particularly surprising given that his disqualification occurred only last August.
Letsrun is a website that is part message board and part news.  The message board,
which functions similarly to Reddit is very active during important running
events, which include during the finals of major events such as
the Olympics and the World Track and Field Championships.  However, it is the
news section of the website that generated articles such as "Was Devon Allen
Screwed? There's At Least a 99\% That He Was" and "The Data Keeps Pouring In and
It Continues to Look Bad For World Athletics and Great For Devon Allen".  Robert
Johnson, who created and runs the Letsrun website argued that there was an error
with the timing equipment at the 2022 Track and Field World Championship and
cited graphs, statistics such as median reaction time, and compared the reaction
times at the 2022 Track and Field World Championship.

In 2009, the International Association of Athletics Federations (Also referred to
as IAAF or World Athletics), which is the governing board for international running
competitions, commissioned a study into changing the reaction time barrier.
The paper analyzed sprinters reaction times and the biological processes 
that went into reacting to the starting gun and concluded 
that athletes can react in as low as 0.08 seconds.  One study by Pain looked at
reaction times for nine male athletes and analyzed their reaction times.  The
fastest athlete of the nine, had on average (even with their two fastest time 
removed) a reaction time of 0.087 seconds and a standard deviation of 0.004
seconds.  Two other athletes were able to react similarly when under certain
muscular tightening conditions \citep{pain2007sprint}.
to IAAF that the disqualification barrier be decreased to allow sprinters to
react faster, and suggested 0.08 or 0.085 seconds as possible new thresholds.
Additionally, it was recommended to IAAF that high speed cameras be used to
change the reaction time criteria from pushing off on the block to body movement.
This could provide a more accurate representation of reaction time, and thus the
possibility of a false disqualification would be lowered \citep{Komi}.  
  A
follow-up study from 2012 by Pilianidis, Mantzouranis, and Kasabalis found
significant differences between reaction times at several world championships
between 1997 and 2009 for the men's 110 meter hurdles.  Another important
conclusion was that for reaction times did not decrease over the twelve years
covered by the research in the study \citep{pilianidis2012start}.  This will
be important when considering the results in section \ref{sec:Results}.

@@

@@

The test statistic of the DS method of the Wilcoxon rank-sum statistic is:
$Z = \frac{S - E(S)}{\sqrt{\hat{\text{Var}}(S)}}$ where 
$S = \mathbb{E}[W^* | X, \delta]$,
$W^* = \frac{1}{N} + \sum_{i=1}^{N} \delta_{i}^{*} R_{i}^{*}$, $R_{i}^{*}$ is
the rank of $X_{i}^{*}$.
The asymptomatic distribution of the Wilcoxon rank-sum statistic was found by
\citet{datta2005rank} and the R package "clusrank" allows for its implementation.
\citep{jiang2017wilcoxon}.

>>>>

\usepackage{amsmath, amsfonts}
\usepackage{booktabs}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{cleveref}
\hypersetup{colorlinks = true, linkcolor = blue, citecolor=blue, urlcolor = blue}
\usepackage{natbib}
\usepackage{float}
\usepackage{setspace}
\usepackage{pdfpages}
\usepackage[pagewise]{lineno}
\usepackage{mwe}
\usepackage{comment}
%\linenumbers*[1]
% %% patches to make lineno work better with amsmath
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
 \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
 \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
 \renewenvironment{#1}%
 {\linenomath\csname old#1\endcsname}%
 {\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
 \patchAmsMathEnvironmentForLineno{#1}%
 \patchAmsMathEnvironmentForLineno{#1*}}%

@@

2)
Outside the blog posts from LetsRun.com, there have not been any published
papers regarding Devon Allen's disqualification in particular.  This is not
particularly surprising given that his disqualification occurred only last August.
Letsrun is a website that is part message board and part news.  The message board,
which functions similarly to Reddit is very active during important running
events, which include during the finals of major events such as
the Olympics and the World Track and Field Championships.  However, it is the
news section of the website that generated articles such as "Was Devon Allen
Screwed? There's At Least a 99\% That He Was" and "The Data Keeps Pouring In and
It Continues to Look Bad For World Athletics and Great For Devon Allen".  Robert
Johnson, who created and runs the Letsrun website argued that there was an error
with the timing equipment at the 2022 Track and Field World Championship and
cited graphs, statistics such as median reaction time, and compared the reaction
times at the 2022 Track and Field World Championship.

3)
In 2009, the International Association of Athletics Federations (Also referred to
as IAAF or World Athletics), which is the governing board for international running
competitions, commissioned a study into changing the reaction time barrier.
The paper analyzed sprinters reaction times and the biological processes 
that went into reacting to the starting gun and concluded 
that athletes can react in as low as 0.08 seconds.  One study by Pain looked at
reaction times for nine male athletes and analyzed their reaction times.  The
fastest athlete of the nine, had on average (even with their two fastest time 
removed) a reaction time of 0.087 seconds and a standard deviation of 0.004
seconds.  Two other athletes were able to react similarly when under certain
muscular tightening conditions \citep{pain2007sprint}.
to IAAF that the disqualification barrier be decreased to allow sprinters to
react faster, and suggested 0.08 or 0.085 seconds as possible new thresholds.
Additionally, it was recommended to IAAF that high speed cameras be used to
change the reaction time criteria from pushing off on the block to body movement.
This could provide a more accurate representation of reaction time, and thus the
possibility of a false disqualification would be lowered \citep{komi2009iaaf}.  
A follow-up study from 2012 by Pilianidis, Mantzouranis, and Kasabalis found
significant differences between reaction times at several world championships
between 1997 and 2009 for the men's 110 meter hurdles.  Another important
conclusion was that for reaction times did not decrease over the twelve years
covered by the research in the study \citep{pilianidis2012start}.  This will
be important when considering the results in section \ref{sec:Results}.

@@

  \begin{equation}
    Y_{ij} = \beta_0 + \beta_1 X_{ij} + \gamma_i + \epsilon_{ij}
    \end{equation}

    where $Y_{ij}$ is the response variable for observation $j$ in year $i$, 
    $X_{ij}$ is the predictor variable for observation $j$ in year $i$, 
    $\beta_0$ is the intercept, $\beta_1$ is the coefficient for $X_{ij}$, 
    $\gamma_i$ is the random effect term for year $i$, and $\epsilon_{ij}$ 
    is the error term for observation $j$ in year $i$.

@@

The test statistic of the DS method of the Wilcoxon rank-sum statistic is:
$Z = \frac{S - E(S)}{\sqrt{\hat{\text{Var}}(S)}}$ where 
$W^* = \frac{1}{N} + \sum_{i=1}^{N} \delta_{i}^{*} R_{i}^{*}$, $R_{i}^{*}$ is
the rank of $X_{i}^{*}$.
The asymptomatic distribution of the Wilcoxon rank-sum statistic was found by
\citet{datta2005rank} and the R package "clusrank" allows for its implementation.
\citep{jiang2017wilcoxon}.
